---
title             : "The title"
shorttitle        : "CTT visuals"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Diego Figueiras"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Montclair State University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "articles.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(psych)
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```



# Methods
We used the formulas presented by @kulas2017approximate. 

Study 2 simulates a bunch of test data and then we generate ICCs based on the IRT model and then we compare that to our CTT estimates. 
## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.



```{r, fig.cap="Relationship between IRT a parameter and CTT corrected-item total correlations."}
g<-abs(qnorm(.5))
g<-0
r2<-.3
ahat<-function(r2){
  r<-(((2.71828)^r2)-(2.71828)^-r2)/(2.71828-(2.71828)^r2)
  ((0.51+(0.02*g)+(0.301*g^2))*r)
  
}

curve(ahat, from=0, to=1, ylim=c(0, 10))
g<-abs(qnorm(.75))
r2<-.7
curve(ahat, lty="longdash",add=TRUE )
g<-abs(qnorm(.1))
r2<-.1
curve(ahat, lty="dotted", add=TRUE)

```
```{r, }
pseudob<-abs(qnorm(.5))

ahat<-function(x){
  r<-(((2.71828)^x)-(2.71828)^-x)/(2.71828-(2.71828)^x)
  ((0.51+(0.02*pseudob)+(0.301*pseudob^2))*r)
  
}
pseudoa<-ahat(.3)
c <- 0

eq <- function(x){c + ((1-c)*(1/(1+2.71828^(-1.7*(pseudoa*(x-pseudob))))))}

curve(eq, from=0, to=1, ylim=c(0, 1))
pseudob<-abs(qnorm(.2))
pseudoa<-ahat(.7)
curve(eq, col="red", add=TRUE)

```

# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
